---
title: "wordle_KB"
output: html_document
date: "2024-02-12"
editor_options: 
  chunk_output_type: console
---

# Creating a Wordle Game

The rules of Wordle are simple: A player has SIX tries to guess a 5-letter word that has been selected at random from a list of possible words. Guesses need to be valid 5-letter words. After each guess, a player receives feedback about how close their guess was to the word, which provides information they can use to try to solve the puzzle. This feedback lets you know whether each letter your guess is either [1] in the solution word and in the correct spot, [2] in the solution word but in the wrong position, or [3] not in the solution word. In app/browser versions of Wordle, this feedback is provided visually using colors, but that need not be the case.

## Step 1

-   Created a custom function called load_dictionary() that takes a single argument, “filename”, that can be used to read in either of the two data files your downloaded.

-   Used that function to create two variables, solution_list and valid_list, that, respectively contain vectors of possible solution words and valid words to guess.

```{r}
# Load files. These files were pulled from our own repository but originally downloaded from https://github.com/difiore/ada-2024-datasets
valid_file <- "https://raw.githubusercontent.com/ecarlson5683/exercise-04/main/collins-scrabble-words-2019.txt"
solution_file <- "https://raw.githubusercontent.com/ecarlson5683/exercise-04/main/google-10000-english-usa-no-swears.txt"


# Creating custom function "load_library"
load_library <- function(filename){
  library(tidyverse)
  read.table(filename, header = TRUE) %>% # create a tibble (data frame)
    pull(words) # ensure that only "words" is being pulled
}

valid_list <- load_library(valid_file)
solution_list <- load_library(solution_file)

str(valid_list)
str(solution_list)
```

## Step 2

-   Winnowed the variable solution_list to only include words that are included in valid_list.
-   Used R help to look at the documentation for the intersection() function to see if it would work.
-   `How many words are in your updated solution_list vector?`

```{r Number of words no_words}
True_solution_list <- intersect(valid_list, solution_list)
str(True_solution_list)
no_words <- length(True_solution_list)
print(paste("There are", no_words, "words"))
```

## Step 3

-   Wrote a custom function called pick_solution() that [1] removes all words from solution_list that are not 5 letters in length, [2] then randomly chooses a single word from those that remain, and [3] then splits that word into a vector of single-character elements. solution_list vector can pass as the argument to the function.
-   Once function worked, assigned the result to a variable called solution.

\*As a bonus, you might include a second argument for your pick_solution() function called “word_length” that makes your function flexible enough to select a solution word that is something other than 5 characters long.

```{r Pick Solution Function}
pick_solution <- function(word){
  subset(word, nchar(word) == 5) |> # Keep only 5 letter long words all words
    sample(1) |>
    str_split_1( "") # randomly select a 5-letter word and make it 5 distinct characters
}
```

```{r Re/Pick Solution}
# Run Code Chunk to Update Solution
solution <- pick_solution(True_solution_list)
print(solution)
```

## Step 4

-   Created two more functions. The first should be called play_wordle() and it should take three arguments: [1] the answer to the puzzle (the value of your solution variable), [2] a list of valid guesses (the contents of your valid_list variable), and [3] a value for “number of guesses”, which you should set to the original Wordle game default of 6.
- Compare the guess to the solution word and generate the necessary feedback, e.g., * for in the word and in the correct position, + for in the word but in the wrong position, and - for not in the word. For this step, wrote a separate “helper” function, evaluate_guess(), called from within play_wordle(). This function should take, as arguments, the player’s guess and the value of the solution variable. 

```{r}
play_wordle_KB <- function(solution, valid_list, num_guesses=6){
  
  feedback <- list() # Create an empty list for results of evaluate_guess()
  
  for (i in 1:num_guesses) { # Create a loop for guessing (replaced `num_guesses` with `i` to count up)
  
    # Tell the player the rules of the game  
    n_length <- length(solution) # Return the length of the solution. This allows some flexibility if you wanted to play a longer- or shorter-character Wordle game
    chances <- if(i == 1){num_guesses # Update chances as player guesses
    } else {num_guesses - (i - 1)}
    
    # Display number of remaining guesses
  print(paste("You now have", (7-i), "chances to guess the", n_length, "lettered word")) # Numbers of guesses left
  
 # Display a key for feedback
  if(i==1){print("Key: * = wrong place in word, - = not in word.")
  } else {NULL}
  
   # Display letters left
  letters_left <- if(i == 1){LETTERS # Update letters as player guesses
  } else {letters_left}
  print(paste("Letters left:", paste(letters_left, collapse = " "))) # Create a string of alphabet then output
   # Prompt player for a guess
  guess <- readline(prompt = paste("Enter guess number", i, "then press <enter>: ")) # Take a character string as an argument and provide a "prompt" entering a line of numeric or character data
  
  # Convert guess entry to uppercase
  guess <- toupper(guess)
  
  # Check validity of guess
  while (is.character(guess)) {
  if(nchar(guess, type = "chars") != 5) { # Check for non-5-letter words and reprompt guess
    cat("Invalid entry: Not a 5-letter word. Please try again. \nLetters left:", letters_left) #\n breaks new line
    guess <- readline(prompt = paste("Enter guess number", i, "then press <enter>: "))
    guess <- toupper(guess)
  } else if(!(guess %in% valid_list)) { # Check for non-valid words and reprompt guess
    cat("Invalid entry: Not a valid word. Please try again. \nLetters left:", letters_left)
    guess <- readline(prompt = paste("Enter guess number", i, "then press <enter>: "))
    guess <- toupper(guess)
  } else { 
    feedback[[i]] <- evaluate_guess(guess, solution) # Name the results of evaluate_guess() to record feedback
    break # End loop so run evaluate_guess() is run only once
  }
  }
  
  # Update the list of letters guessed
  guess_letters <- str_split(guess, "")[[1]]
  letters_left <- setdiff(letters_left, guess_letters) # Overwrite letters_left to remove letters of guess
  
  # Create a matrix of feedback for game summary
  summary <- matrix(unlist(feedback, use.names = FALSE), ncol = 5, byrow = TRUE)
  dimnames(summary) <-list(rep("", dim(summary)[1]), rep("", dim(summary)[2]))
  
  # Check if puzzle is solved
  if(identical(guess_letters, solution)) {
    cat("\nCongratulations, you won!\n\nGuess summary:\n")
    print(noquote(summary)) # Print guess and feedback history without quotes
    break # end loop
  } else if (!identical(guess_letters, solution) & i < num_guesses) { 
    # Loop to beginning, now with updated letter list for remaining guesses
  } else {
    cat("\nNo more guesses. Better luck next time! Solution:", solution, "\n\nGuess summary:\n") # Loop ends, notify player of loss and correct solution
    print(noquote(summary)) # Print guess and feedback history without quotes
  }
  
  }
  
}
  
  
```

```{r Helper Function}
# Helper function to evaluate valid guesses

evaluate_guess <- function(guess, solution){
  
  library(tidyverse)
  
  guess_letters <- str_split(guess, "")[[1]]
  
  for (i in 1:length(guess_letters)) {
  output <- case_when(guess_letters[] == solution[] ~ guess_letters[], # Fill in correct letter
                      guess_letters[] %in% solution[] ~ "*",
                      guess_letters[] != solution[] ~ "-"
                      )
  }
  print(paste("", guess_letters, "\n", output, "\n")) # Show guess, then feedback, then move to a new line
  feedback <- list(guess = guess_letters, output = output) # Create list of objects to be exported from function
  return(feedback) # Export from function to store feedback
}
```

```{r Let's PLay!}
 play_wordle_KB(solution, valid_list)
```


